---
title: "Bharatviz demo"
format: html
---

BharatViz api allows you to generate publication-ready India choropleth maps with minimal code (3 lines):

```r
library(R6)
source("bharatviz.R")
bv <- BharatViz$new()
bv$save_all_formats(your_data, basename = "publication_map")
```

---

Disclaimer: The notebook takes you through different uses cases of the API. Please note that all the values used in this demo are arbitrary. These values are very likely incorrect - they are only used for demonstrating the capability of BharatViz API and should not be used for drawing any insights.

---

## Installation

```{r}
# Install required packages if not already installed
if (!require("R6")) install.packages("R6")
if (!require("httr")) install.packages("httr")
if (!require("jsonlite")) install.packages("jsonlite")
if (!require("base64enc")) install.packages("base64enc")
if (!require("png")) install.packages("png")

library(R6)
library(httr)
library(jsonlite)
library(base64enc)
library(png)

# Download the R client if not present
if (!file.exists("bharatviz.R")) {
  download.file(
    "https://gist.githubusercontent.com/saketkc/7b227151bde59dfa31fd2b1dd15f0c67/raw/d13c9cdc7dd814a6be9d53483e4e48e3db6db9db/bhartviz.R",
    "bharatviz.R",
    quiet = TRUE
  )
}
source("bharatviz.R")

API_URL <- "https://bharatviz.saketlab.org"

# Initialize client
bv <- BharatViz$new(api_url = API_URL)
```

---

# State level maps

## Quickest way - from dictionary

```{r}
# Create data from a simple dictionary
literacy_data <- list(
  "Kerala" = 93.9,
  "Delhi" = 86.3,
  "Maharashtra" = 82.9,
  "Tamil Nadu" = 80.3,
  "Gujarat" = 79.3,
  "Karnataka" = 75.6,
  "Uttar Pradesh" = 69.7,
  "Bihar" = 63.8,
  "Rajasthan" = 67.1
)

# Convert and display with one line!
quick_map(
  from_dict(literacy_data),
  title = "Literacy rate by state",
  legend_title = "Percentage (%)",
  color_scale = "greens"
)
```

## From pandas dataFrame

```{r}
# Create dataframe with state data
state_data <- data.frame(
  state = c(
    "Maharashtra",
    "Kerala",
    "Karnataka",
    "Tamil Nadu",
    "Gujarat",
    "Rajasthan",
    "Uttar Pradesh",
    "West Bengal",
    "Andhra Pradesh",
    "Telangana"
  ),
  gdp_growth = c(7.8, 6.5, 8.3, 8.7, 8.1, 7.3, 6.8, 7.6, 7.8, 8.2)
)

print(head(state_data))

# Generate map - dataframe is auto-ingested!
result <- bv$generate_map(
  state_data,
  title = "State gdp growth rate (2023)",
  legend_title = "% Growth",
  color_scale = "blues"
)
bv$show_map(result)
```

## Save to png/svg/pdf

```{r eval=FALSE}
# Save all formats at once: PNG, SVG, PDF
save_all_formats(
  bv,
  state_data,
  basename = "state_gdp_growth",
  title = "State gdp growth rate",
  legend_title = "% Growth",
  color_scale = "viridis"
)
```

## Custom column names

```{r}
# Dataframe with custom column names
custom_df <- data.frame(
  state_name = c(
    "Punjab",
    "Haryana",
    "Himachal Pradesh",
    "Uttarakhand",
    "Jharkhand"
  ),
  unemployment_rate = c(5.2, 4.8, 3.9, 4.5, 6.1)
)

# Convert using helper function
data <- from_dataframe(
  custom_df,
  state_col = "state_name",
  value_col = "unemployment_rate"
)

# Generate map
result <- bv$generate_map(
  data,
  title = "Unemployment rate by state",
  legend_title = "Percentage",
  color_scale = "reds"
)
bv$show_map(result)
```

## Compare color scales

```{r}
# Compare multiple color scales side-by-side
compare_scales(
  bv,
  state_data,
  scales = c("spectral", "viridis", "plasma", "blues", "reds", "greens")
)
```

---

# District-level maps

Currently, BharatViz supports three different district maps:

- **[LGD](https://lgdirectory.gov.in/)**: Latest district boundaries (default)
- **NFHS5**: NFHS-5 survey boundaries
- **NFHS4**: NFHS-4 survey boundaries

## Quick district map

```{r}
# Sample district data
district_data <- list(
  list(state = "Maharashtra", district = "Mumbai", value = 89.7),
  list(state = "Maharashtra", district = "Mumbai Suburban", value = 90.9),
  list(state = "Maharashtra", district = "Pune", value = 86.2),
  list(state = "Maharashtra", district = "Nagpur", value = 89.5),
  list(state = "Maharashtra", district = "Thane", value = 84.5),
  list(state = "Karnataka", district = "Bengaluru Urban", value = 88.7),
  list(state = "Karnataka", district = "Mysuru", value = 73.4),
  list(state = "Karnataka", district = "Dharwad", value = 80.0),
  list(state = "Kerala", district = "Thiruvananthapuram", value = 93.0),
  list(state = "Kerala", district = "Kottayam", value = 97.2),
  list(state = "Kerala", district = "Ernakulam", value = 95.7),
  list(state = "Tamil Nadu", district = "Chennai", value = 90.2),
  list(state = "Tamil Nadu", district = "Coimbatore", value = 84.2),
  list(state = "Delhi", district = "Central", value = 86.3),
  list(state = "Delhi", district = "South", value = 87.4)
)

# Display using quick function
quick_districts_map(
  district_data,
  title = "District literacy rates",
  legend_title = "Literacy %",
  color_scale = "spectral",
  map_type = "LGD"
)
```

## Districts from dataframe

```{r}
# Create district dataframe
districts_df <- data.frame(
  state = c(
    "Maharashtra",
    "Maharashtra",
    "Karnataka",
    "Karnataka",
    "Kerala",
    "Kerala"
  ),
  district = c(
    "Mumbai",
    "Pune",
    "Bengaluru Urban",
    "Mysuru",
    "Thiruvananthapuram",
    "Kottayam"
  ),
  value = c(89.7, 86.2, 88.7, 73.4, 93.0, 97.2)
)

print(districts_df)

# Generate districts map - dataframe is auto-converted!
result <- bv$generate_districts_map(
  districts_df,
  map_type = "LGD",
  title = "Sample district data",
  legend_title = "Values",
  color_scale = "viridis",
  show_state_boundaries = TRUE
)
bv$show_map(result)
```

## Different district map types (LGD vs NFHS)

```{r}
# Same data, different boundary definitions (names)
sample_districts <- list(
  list(state = "Maharashtra", district = "Mumbai", value = 89.7),
  list(state = "Karnataka", district = "Bengaluru Urban", value = 88.7),
  list(state = "Kerala", district = "Thiruvananthapuram", value = 93.0),
  list(state = "Tamil Nadu", district = "Chennai", value = 90.2),
  list(state = "Delhi", district = "Central", value = 86.3)
)

result <- bv$generate_districts_map(
  sample_districts,
  map_type = "LGD",
  title = "LGD district boundaries"
)
bv$show_map(result)

# Generate with NFHS5 boundaries
cat("\nGenerating NFHS5 map...\n")
result <- bv$generate_districts_map(
  sample_districts,
  map_type = "NFHS5",
  title = "NFHS-5 district boundaries"
)
bv$show_map(result)
```

## Custom column names for districts

```{r}
# Dataframe with custom column names
custom_districts <- data.frame(
  state_name = c("Maharashtra", "Maharashtra", "Kerala", "Kerala"),
  district_name = c("Mumbai", "Pune", "Thiruvananthapuram", "Kottayam"),
  infant_mortality = c(24.0, 28.0, 8.0, 7.0)
)

# Convert using helper function
data <- from_dataframe_districts(
  custom_districts,
  state_col = "state_name",
  district_col = "district_name",
  value_col = "infant_mortality"
)

# Generate map
result <- bv$generate_districts_map(
  data,
  title = "Infant mortality rate by district",
  legend_title = "Per 1000 births",
  color_scale = "reds",
  invert_colors = TRUE # Lower is better
)
bv$show_map(result)
```

---

# State-District maps

Display districts of a single state with different boundary definitions. Useful for detailed regional analysis.

## Rajasthan with LGD boundaries

```{r}
# Debug: Check if bv is available and has the method
cat("=== DEBUG START ===\n")
cat("Class of bv:", class(bv), "\n")
cat("generate_state_districts_map exists:", "generate_state_districts_map" %in% names(bv), "\n")
cat("=== DEBUG END ===\n")
```

```{r}
# District data for Rajasthan only
rajasthan_districts <- data.frame(
  state = rep("Rajasthan", 5),
  district = c("Jaipur", "Jodhpur", "Udaipur", "Ajmer", "Bikaner"),
  literacy_rate = c(75.5, 68.2, 71.3, 70.1, 67.9)
)

# Generate state-district map using LGD boundaries
result <- bv$generate_state_districts_map(
  rajasthan_districts,
  state = "Rajasthan",
  map_type = "LGD",
  title = "Rajasthan: Literacy rate by district (LGD)",
  legend_title = "Literacy %",
  color_scale = "blues"
)
bv$show_map(result)
```

## Rajasthan with NFHS5 boundaries

```{r}
# Same district data with NFHS5 boundaries

result <- bv$generate_state_districts_map(
  rajasthan_districts,
  state = "Rajasthan",
  map_type = "NFHS5",
  title = "Rajasthan: Literacy rate by district (NFHS5)",
  legend_title = "Literacy %",
  color_scale = "blues"
)
bv$show_map(result)
# Code reference
```

## Rajasthan with NFHS4 boundaries

```{r}
# Same data with NFHS4 boundaries
result <- bv$generate_state_districts_map(
  rajasthan_districts,
  state = "Rajasthan",
  map_type = "NFHS4",
  title = "Rajasthan: Literacy rate by district (NFHS4)",
  legend_title = "Literacy %",
  color_scale = "blues"
)
bv$show_map(result)
```

## Compare different map types for same state

```{r}
# how boundaries changed between surveys
result_lgd <- bv$generate_state_districts_map(
  rajasthan_districts,
  state = "Rajasthan",
  map_type = "LGD",
  title = "LGD Boundaries"
)
bv$show_map(result_lgd)

# NFHS5
result_nfhs5 <- bv$generate_state_districts_map(
  rajasthan_districts,
  state = "Rajasthan",
  map_type = "NFHS5",
  title = "NFHS5 Boundaries"
)
bv$show_map(result_nfhs5)

# NFHS4
result_nfhs4 <- bv$generate_state_districts_map(
  rajasthan_districts,
  state = "Rajasthan",
  map_type = "NFHS4",
  title = "NFHS4 Boundaries"
)
bv$show_map(result_nfhs4)
```

---

# Advanced features

## Get metadata

```{r}
# Generate map and get metadata
result <- bv$generate_map(
  state_data,
  title = "Gdp growth with metadata",
  formats = c("png", "svg", "pdf")
)

# Access metadata
cat("Map statistics:\n")
cat(sprintf("  Min value: %s\n", result$metadata$minValue))
cat(sprintf("  Max value: %s\n", result$metadata$maxValue))
cat(sprintf("  Mean value: %.2f\n", result$metadata$meanValue))
cat(sprintf("  Color scale: %s\n", result$metadata$colorScale))
```

## Inverted color scales

```{r}
# For metrics where lower is better (e.g., unemployment, mortality)
unemployment_data <- list(
  "Kerala" = 4.2,
  "Maharashtra" = 4.8,
  "Karnataka" = 3.9,
  "Bihar" = 7.5,
  "Uttar Pradesh" = 6.1
)

result <- bv$generate_map(
  from_dict(unemployment_data),
  title = "Unemployment rate (Inverted scale)",
  legend_title = "Percentage",
  color_scale = "reds",
  invert_colors = TRUE # Red for high unemployment
)
bv$show_map(result)
```

##  Hide labels for cleaner maps

```{r}
# Generate map without labels
result <- bv$generate_map(
  state_data,
  title = "Clean map (No labels)",
  color_scale = "blues",
  hide_state_names = TRUE,
  hide_values = TRUE
)
bv$show_map(result)
```

---

# Quick Reference


### States
```r
# From dictionary - quickest!
quick_map(from_dict(list("Kerala" = 93.9)), title = "My map")

# From dataframe
result <- bv$generate_map(df)
bv$show_map(result)

# Save pdf
save_all_formats(bv, data, basename = "my_map")
```

### Districts
```r
# From list
quick_districts_map(data, title = "Districts", map_type = "LGD")

# From dataframe
result <- bv$generate_districts_map(df, map_type = "LGD")
bv$show_map(result)

# Save pdf
save_all_formats_districts(bv, data, basename = "districts")
```

### State-District (Single State)
```r
# Districts of a single state
result <- bv$generate_state_districts_map(
  df,
  state = "Rajasthan",
  map_type = "LGD"
)
bv$show_map(result)

# Save pdf
save_all_formats_state_districts(bv, data, state = "Rajasthan", basename = "rajasthan")
```

## Available color scales

**Sequential:** `blues`, `greens`, `reds`, `oranges`, `purples`, `pinks`, `viridis`, `plasma`, `inferno`, `magma`

**Diverging:** `spectral`, `rdylbu`, `rdylgn`, `brbg`, `piyg`, `puor`

## Map types (Districts)

- **LGD**: Latest Government of India district boundaries
- **NFHS5**: National Family Health Survey 5 (2019-21)
- **NFHS4**: National Family Health Survey 4 (2015-16)
- **Census**: Census of India 1941-2011 boundaries

---
